---
title: "Task 3 and 4"
author: "Jenna Wisniewski"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
```

## Read in the Urchin Data
```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```
## Sectiom 1- Size Comparison: Purple and Red Urchins

### Exploratory Data Visualization
In this section I look at the distribution of sizes between red and purple urchins using two types of graphs. 

#### Create a Faceted Histogram
```{r}
ggplot(data = urchins, aes(x = size))+
  geom_histogram(bins = 15) +
  facet_wrap(~common_name)+
   labs(x = "Size of Urchin", 
       y = "Number of Urchins")


```

#### Create a Boxplot
```{r}
ggplot(data = urchins, aes(x = common_name, y = size))+
  geom_boxplot() +
   labs(x = "Type of Urchins", 
       y = "Size Distribution of Urchins")
```

### Descriptive Statistics Table
In this section I will create and look at the statistics of for both red urchin and purple urchin populations (mean, median, standard deviation, n)
```{r}
urchin_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(
    mean_size = mean(size),
    median_size = median(size),
    sd_size = sd(size),
    n = n())

   

urchin_stats 
```
### Overall Takeaways 
When looking at this data, 

## Section 2- Exploring Purple Urchin Counts by Site and Year

### Data Wrangling
In this section I will create a subset of the data that 
-filters observations for only purple urchins
-Converts the date to class "Date"
-Create a new column that only contains the year
-Create a new column that contains the full name of each of the five sites

```{r}
 purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
site == "NAPL" ~ "Naples Reef",
site == "AQUE" ~ "Arroyo Quemado Reef",
site == "CARP" ~ "Carpinteria Reef",
site == "MOHK" ~ "Mohawk Reef",
site == "IVEE" ~ "Isla Vista Reed"
))

         

```
### Descriptive Statistics Table 
In this section I will create a table containing the total number of purple urchins observed by site and year

```{r}
urchin_count <- purple_urchins %>% 
  select("year", "site_full") %>% 
   count(year, site_full) 

urchin_count

```

### Make an Explanantory Graph
In this section I will create an exploratory graph of the purple urchins recorded at each site over time for the years included in the study. 
```{r}
ggplot(data = urchin_count, aes(x = year, y = n))+
  geom_col(aes(color = site_full))+
  facet_wrap(~site_full)+
  labs(x = "Year", 
       y = "Total Number of Purple Urchins")
  
```


