---
title: "Task 4"
author: "Jenna Wisniewski"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
```

## Read in the Urchin Data
```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Compare the mean size of purple urchins with no giant kelp removal and with annual giant kelp removal in sections of Naples Reef in Jabuary 2010

### Wrangle the Data
In this section I will wrangle the urchins data to 
-only include purple urchins
-only include observations from Naples Reef
-only keep observations for January 2010

```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date)) %>% 
  mutate(site_full = case_when(
site == "NAPL" ~ "Naples Reef",
site == "AQUE" ~ "Arroyo Quemado Reef",
site == "CARP" ~ "Carpinteria Reef",
site == "MOHK" ~ "Mohawk Reef",
site == "IVEE" ~ "Isla Vista Reed"
))

napl_urchins <- purple_urchins %>% 
  filter(site_full == "Naples Reef") %>% 
  filter(year == 2010, month == 1)

```

### Create Exploratory Graphs
In this section I will create two exploratory graphs in order to look at population distribution stemming from the two different treatments (no giant kelp removal, and annual giant kelp removal)

### Creat a histogram
```{r}
ggplot(data = napl_urchins, aes(x = size))+
  geom_histogram(bins = 15) +
  facet_wrap(~treatment)+
   labs(x = "Size of Urchin(cm)", 
       y = "Number of Urchins of that Size")
```
#### Create a QQPlot


```{r}
ggplot(data = napl_urchins, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~treatment)
```
### Main Takeaways
My main takeaways from these graphs are

### Create a Basic Statistic Summary Table 
```{r}
np_urchin_stats <- napl_urchins %>% 
  group_by(treatment) %>% 
  summarize(
    mean_size = mean(size),
    median_size = median(size),
    sd_size = sd(size),
    n = n())

np_urchin_stats
```

### Perform a two-sample t-test
In this section I perform a two-sample t-test to compare means between the two groups of treatments
#### Data Wrangling
Split the two treatments into two different vectors
```{r}

annual_sample <- napl_urchins %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_sample <- napl_urchins %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)
```
### Perfrom a two-sample t-test
```{r}
t.test(annual_sample, control_sample)
```


